---
title: "p8105_hw2_qz2492"
author: "qz2492"
date: "2022-10-02"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

```{r read and clean data, message=FALSE}
# Read data and clean its names
NYC_T = read.csv('./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>%
  janitor::clean_names()

# Check which columns got NULL values
NYC_T %>% naniar::miss_var_summary()

# Convert NULL values into chr to make route 8~11 consistent with others
NYC_T[is.na(NYC_T)] = '0'

# Select the columns we want and convert entry variable from chr to logical
NYC_T = select(NYC_T,
  line, station_name, station_latitude, station_longitude,
  starts_with("route"), entry, vending, entrance_type, ada
) %>% mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

This dataset mainly contains names and locations of every station, which lines are they in, 
whether they have vending machines and equipment for disabled. My cleaning steps include reading and cleaning names, processing NULL values and selection for columns. The resulting dataset has 1868 rows x 19 columns. And it is not tidy because of the redundance in `route` variable.

```{r distinct stations, eval=FALSE, message=FALSE, include=TRUE}
NYC_T %>% 
  select(station_name, line) %>% 
  distinct
```

According to the code and output, there are 465 distinct stations.

```{r ADA compliance, eval=FALSE, message=FALSE, include=TRUE}
NYC_T %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

According to the code and output, there are 84 ADA compliant stations.

```{r compute proportions}
NYC_T %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

According to the code and output, there are 37.7% station entrances / exits without vending allow entrance.

```{r distinct stations serve the A train, eval=FALSE, message=FALSE, include=TRUE}
NYC_T %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct
```
 
 According to the code and output, there are 60 distinct stations serving the A train.

```{r proportions that ADA compliant, eval=FALSE, message=FALSE, include=TRUE}
NYC_T %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

According to the code and output, there are 17 distinct stations serving the A train and ADA compliant.

## Problem 2

```{r read and clean the sheet for Mr. Trash Wheel, warning=FALSE}
# Read data and omit non-data entries
MrTW = read_excel('./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx', sheet = "Mr. Trash Wheel", range = "A2:N535") %>% janitor::clean_names()

# omit rows that do not include dumpster-specific data
MrTW$dumpster = as.numeric(MrTW$dumpster)
MrTW = MrTW[!is.na(MrTW$dumpster),]

# round the number of sports balls
MrTW$sports_balls = as.integer(MrTW$sports_balls)
```


```{r read and clean the sheet for Professor Trash Wheel, warning=FALSE}
# Read data and omit non-data entries
ProfTW = read_excel('./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx', sheet = "Professor Trash Wheel", range = "A2:N117") %>% janitor::clean_names()

# omit rows that do not include dumpster-specific data
ProfTW$dumpster = as.numeric(ProfTW$dumpster)
ProfTW = ProfTW[!is.na(ProfTW$dumpster),]

# round the number of sports balls
ProfTW$sports_balls = as.integer(ProfTW$sports_balls)
```

```{r combine two datasets to produce a tidy one}
# add a column in each tibble to identify them
MrTW = add_column(MrTW, sheet = 'Mr. Trash Wheel', .before = 1)
ProfTW = add_column(ProfTW, sheet = 'Professor Trash Wheel', .before = 1)

# combining
CombinedTW = bind_rows(MrTW, ProfTW)
```

There are `r nrow(MrTW)` rows, `r ncol(MrTW)` columns and `r nrow(MrTW)* ncol(MrTW)` observations in _Mr. Trash Wheel data set_,
`r nrow(ProfTW)` rows, `r ncol(ProfTW)` columns and `r nrow(ProfTW)* ncol(ProfTW)` observations in _Professor Trash Wheel data set_, and
`r nrow(CombinedTW)` rows, `r ncol(CombinedTW)` columns and `r nrow(CombinedTW)* ncol(CombinedTW)` observations in _Combined data set_.
All of them have the same columns, `sheet` indicating which sheets they belong to, `Dumpster` is the dumpster number, `month`, `year`, `date` indicate the time when these dumpster worked, `weight_tons` and `volume_cubic_yards` describe how much trash each dumpster removed. For  _Mr. Trash Wheel data set_, the _weight_ is from `r min(MrTW$weight_tons)` to `r max(MrTW$weight_tons)` tons, with the mean of `r round(mean(MrTW$weight_tons), digits = 2)`  tons. For  _Professor Trash Wheel data set_, the _weight_ is from `r min(ProfTW$weight_tons)` to `r max(ProfTW$weight_tons)` tons, with the mean of `r round(mean(ProfTW$weight_tons), digits = 2)` tons.  
For this data,  the total weight of trash collected by Professor Trash Wheel is `r sum(ProfTW$weight_tons)` tons, the total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(MrTW %>% filter(year == 2020) %>% select(sports_balls))`.
